Distribution-oblivious Programming with Dish

### 2019-11-20 pldi/draft.tex
Complete a first draft of Section 5

In order to effectively apply optimizations and distribute a shell
script, it first has to be translated to a manipulable
representation. As mentioned in \ref{}, the two main abstractions of
the shell, are files that contain data, and commands that communicate
through these files. We propose a slight variation of the standard
dataflow graph, that is commonly found in stream processing systems,
as the representation of a distributed shell computation. The dataflow
graph nodes represent commands, and the edges represent files.

\kk{Before that I have to mention that the basic data type $D$ is a
  line of characters.}

@@

Edges in the dataflow graph represent files, the basic data
abstraction of the shell. They are used as communication channels
between nodes in the graph, and as the input or output of the whole
graph. Edges are represented as possibly unbounded streams of type
$D*$, where $D$ represents a line, that is a sequence of characters
followed by \verb|\n| \kk{or EOF?}.

\kk{The following paragraph falls in the trap of just explaining
  without motivating}

Edges could either refer to a specific named file in the file system,
or just be FIFO pipes that used for interprocess communication \kk{Are
  there any other file types, such as URLs, ...?}. Edges that refer to
real files, contain their exact number of lines, while edges that are
fifo pipes can have an upper bound on the number of lines that they
will transfer through them, thus being bounded in size.

\kk{Depending on which of the two paragraphs goes last (nodes/edges),
  I have to explain there that the last incoming edge to a node must
  be unbounded (if it doesn't point to a resource).}

\kk{Text point: The main shell abstraction is files, which are
  sequences of lines, where the order matters. So we want a
  representation that captures order of inputs and outputs (and ours
  does because its node in the graph has a sequence of inputs.}

@@

\kk{Mergers and splitters, and how they are implemented using shell constructs (cat and head)}

As we described above, a node could have many incoming edges

\kk{Section notes:}

\kk{Files have an associated number of lines}

\kk{Explain how the assumptions made in the front end give as correct
  distribution}

>>>>

\kk{Try to be more formal in the section, by using the notation that
  is defined in the nodes paragraph.}

In order to effectively apply optimizations and distribute a shell
script, it first has to be translated to a manipulable
representation. As mentioned in \ref{}, the two main abstractions of
the shell, are files that contain data, and commands that communicate
through these files. We propose a slight variation of the standard
dataflow graph, that is commonly found in stream processing systems,
as the representation of a distributed shell computation. The main
benefit of the dataflow graph model is that it clearly exposes
operator level parallelism, as different nodes can be independent
processing workers. In our case each node represents a command, and
each edge represents a file.

\kk{Before that I have to mention that the basic data type $D$ is a
  line of characters. If we move the edges section before the nodes
  that will be fine.}

@@

Edges in the dataflow graph represent files, the basic data
abstraction of the shell. They are used as communication channels
between nodes in the graph, and as the input or output of the whole
graph. Edges are represented as possibly unbounded streams of type
$D*$, where $D$ represents a line, that is a sequence of characters
followed by \verb|\n| \kk{or EOF?}. They could either refer to a
specific named file in the file system, or just be FIFO pipes that
used for interprocess communication \kk{Are there any other file
  types, such as URLs, ...?}.

\kk{Show an example pipe and its graph}.

In order to expose operator parallelism with the output of a command,
files that don't point to a resource can have an upper bound on the
lines that they can tranfer. The following example illustrates why
this is essential.

\kk{An example of a command that has two outputs and one goes to one
  command, while another output goes to a different one. Maybe only
  show the graph here and not the command itself.}

On the other hand, some forms of data parallelism can be exposed when
knowing the size of the input files. As mentioned in \ref{} some pure
commands (such as cat -n) only need line information to become
stateless, and knowing the size of a file could allow the system to
split it in different chunks that can be processed independently. To
account for that, edges that refer to an input resource contain the
number of lines of the file that they refer to. \kk{At the moment this
  is not supported by our real system. The whole paragraph above can
  go if we decide that it is not a good idea to talk about it now.}

Finally, the edges that don't refer to a resource and do not start
from a node in the graph represent the graph input, while the edges
that don't refer to a resource and do not point to a node in the graph
are its outputs.

\kk{(Maybe) Depending on which of the two paragraphs goes last
  (nodes/edges), I have to explain there that the last incoming edge
  to a node must be unbounded (if it doesn't point to a resource). I
  am actually not sure whether I should include this here.}

\kk{(Maybe) It might be beneficial to mention that many inputs do not
  represent the same thing as many outputs. These two notions are not
  symmetric.}

@@

\kk{Mergers and splitters, and how they are implemented using shell
  constructs (cat and head). This subsection might be a better fit in
  some other section, maybe something related to the planner or the
  implementation, as it is a very elegant way to implement splitters
  and mergers while staying completely inside the shell programming
  model.}

It is common in the dataflow graph for a node to have many incoming
edges, or many outgoing edges (when all but the last outgoing edges
are bounded). However some nodes, might read their input from one file
(e.g. stdin). In this case, the distributed implementation needs to
concatenate input files, and split output files. Here is an example:

\kk{Show an example graph! of cat | tr | sort. The command in the
  middle must only take input from one file (maybe just stdin).}

\kk{This transition needs to be stronger} Luckily, the shell offers
primitives for file manipulations that can be used to implement a
merger and a splitter. More concretely a merger can be implemented by
a simple \verb|cat|. Given a set of input files \verb|in1, in2, ...|
and an output file \verb|out|, a merger can be implemented as:

\begin{lstlisting}[language=sh, float=h, numbers=none]
 cat $in1 $in2 ... > $out
\end{lstlisting}

\noindent
On the other hand, given an input file \verb|in| that is a FIFO pipe,
and two output files \verb|out1, out2| the first of which is bounded
to \verb|N| lines, a 2-splitter can be implemented as:

\begin{lstlisting}[language=sh, float=h, numbers=none]
 head -n $N $in > $out1 ; cat $in > $out2
\end{lstlisting}

\noindent
Using the 2-splitter as a building block, a pipe can be split to
arbitrarily many different pipes.

\kk{Do I also need to include a splitter for normal files?}

\kk{Intermediate Representation Section notes:}

\kk{Explain how the assumptions made in the front end give as
  correct/safe operator level parallelism. Maybe this must go in the
  Front-end section.}
